---
title: "Align grids to Bioko Grid"
author: "Brendan F Fries, Carlos A Guerra, et al"
output:
  pdf_document:
    number_sections: true
    df_print: kable
    toc: true
header-includes:
  - \usepackage{palatino}
# The params block means you can put the data anywhere you like and
# then render this document by passing a parameter to the rmarkdown render command.
params:
  data_directory: !r fs::path(rprojroot::find_root(rprojroot::is_rstudio_project), "inst", "extdata")
---

# Make other datasets align to BIMEP study grids
The BIMEP study uses two grids, at 100m and 1km resolution.
This section reads population datasets and regrids them
onto the BIMEP study datasets.


The output is all GeoTIFFs with names of the form data-on-grid.
The data are BIMEP, WorldPop, HRSL, and LandScan. The grids
are

- area, a 1km grid used by BIMEP.
- worldpop, a 100m grid used by WorldPop
- 1kworldpop, a 1km grid that aggregates and aligns with the WorldPop grid.
- landscan, the 1km LandScan grid
- hrsl, the 100m HRSL grid.

## Load BIMEP Early

The BIMEP data will have a "pop" feature for the population in the pixel.
Load it now so that we have its bounding box.

```{r bimep_raster}
bimep_feature <- bimep_population_as_points(local_directory = data_dir)
names(bimep_feature)
```
We see the population as `pop` and the other column, which is the geometry
for each point.

## Align HRSL and BIMEP to BIMEP 1k Grid
The Bioko grids define cells where BIMEP assigns residents of
the island. One is a 1 km grid, the other a 100 m grid.
If we read those grids, we can see that the input files
are projected into UTM 32N, which makes all distances in meters.
That's very convenient, but the original grids were defined as
a regular grid in latitude and longitude.
```{r grids, dependson = c("data_dir", "download")}
grids <- read_bioko_grids(local_directory = data_dir)
grid_info <- parameters_of_km_grid(grids$coarse, unproject = TRUE)
```
You can see from the description above that the coarse
and fine grids are shapefiles. Are these grids complete, or are they partial?
```{r complete_grids}
area_per_pixel <- function(grid) {
  bbox <- st_bbox(grid)
  (bbox["xmax"] - bbox["xmin"]) * (bbox["ymax"] - bbox["ymin"])
}
cat("area per pixel coarse", area_per_pixel(grids$coarse) / 3894, "\n")
cat("area per pixel fine", area_per_pixel(grids$fine) / 197086, "\n")
```
We see that the area per pixel for the coarse is about $10^6\:\mbox{m}^2$, which
means it's probably a complete grid. The fine grid looks far from complete
because we expect $10^4\:\mbox{m}^2$ per pixel.

If we want to make such a grid, we could use
```{r grid_copy, dpeendson = c("download")}
coarse_grid_copy <- make_fresh_grid(grid_info$bbox, grid_info$dimensions)
```
We can show the original and copy in unprojected space
in order to see that they overlap.
```{r compare_1km_grids, fig.cap = "The original and our copy overlap.", dependson = "grid_copy"}
grid_1km <- sf::st_transform(grids$coarse, crs = latitude_longitude_projection)
tm_shape(coarse_grid_copy) +
  tm_polygons(alpha = 0.1, border.col = "red") +
  tm_shape(grid_1km) +
  tm_polygons(alpha = 0.0, border.col = "blue")
```


What if we directly make a raster version of the 1km grid?
```{r make_raster}
km_raster <- raster::raster(
  ncol = grid_info$dimensions[1],
  nrow = grid_info$dimensions[2],
  xmn = grid_info$bbox["xmin"],
  xmx = grid_info$bbox["xmax"],
  ymn = grid_info$bbox["ymin"],
  ymx = grid_info$bbox["ymax"]
  )
raster::projection(km_raster) <- latitude_longitude_projection
```

Let's get some points to project to that raster.
The HRSL gives the location of a roof, or NA values. It doesn't say where
there could be population but isn't. Other population maps assign
a population of 0 to all non-ocean areas, including lakes. LandScan says
it extends the land area by a bit, too. We will use a shapefile for
Bioko in order to assign zero values to the HRSL.
```{r hrsl_plot}
hrsl_raster <- read_hrsl(local_directory = data_dir)
hrsl_raster_crop <- raster::crop(hrsl_raster, bimep_feature, snap = "out")
hrsl_zero_mask <- raster::rasterize(bioko_sf, hrsl_raster_crop, field = 0)
hrsl_raster_zero <- raster::cover(hrsl_raster_crop, hrsl_zero_mask)
hrsl_points_sf <- hrsl_points(hrsl_raster_crop)
# hrsl_points_sf[is.na(hrsl_points_sf[[1]]), 1] <- 0
plot(hrsl_raster_zero)
```


And do the projection using a builtin function of the raster package.
```{r project_to_raster}
km_hrsl_raster <- raster::rasterize(
  hrsl_points_sf,
  km_raster,
  field = "population_gnq_2018.10.01",
  fun = sum
  )
area_zero_raster <- raster::rasterize(bioko_sf, km_hrsl_raster, field = 0)
km_hrsl_raster <- raster::cover(km_hrsl_raster, area_zero_raster)
```

```{r}
plot(km_hrsl_raster)
```

Rasterize that pop feature, summing it per pixel.
```{r}
bimep_raster <- raster::rasterize(
  bimep_feature,
  km_raster,
  field = "pop",
  fun = sum
  )
names(bimep_raster)
```
```{r}
plot(bimep_raster)
```

We will write all raster layers the same way.
```{r}
write_one <- function(layer, basename) {
  raster::writeRaster(
    layer,
    filename = fs::path(data_dir, basename),
    format = "GTiff",
    overwrite = TRUE
    )
}
```

Write these as layers to a GeoTIFF.
```{r write_hrsl_bimep}
names(km_hrsl_raster) <- c("HRSL")
names(bimep_raster) <- c("BIMEP")
write_one(km_hrsl_raster, "hrsl_on_area")
write_one(bimep_raster, "bimep_on_area")
write_one(hrsl_raster_zero, "hrsl_on_hrsl")
```


```{r, eval = FALSE}
grids$coarse["hrsl"] <- hrsl_points_sf[[1]]
tm_shape(grids$coarse[grids$coarse$hrsl > 0,]) +
  tm_polygons("hrsl", title = "HRSL", palette = "Greens", style = "log10") +
  tm_layout(legend.title.size = 1, legend.text.size = 1)
```

## Align BIMEP to LandScan Grid

The LandScan grid is an Arc/Info Binary Grid, so it's the adf type.
It uses pixel sizes of 0.008333 in both dimensions.
It's the whole world, so we will subset it.

```{r landscan}
landscan <- read_landscan(local_directory = data_dir)
landscan <- raster::crop(landscan, bimep_feature, snap = "out")
plot(landscan)
```
Now put the bimep 100m data on the LandScan raster.
```{r}
bimep_landscan_raster <- raster::rasterize(
  bimep_feature,
  landscan,
  field = "pop",
  fun = sum
  )
names(bimep_landscan_raster)
```

```{r}
plot(bimep_landscan_raster)
```

```{r write_landscan_bimep}
names(bimep_landscan_raster) <- c("HRSL")
names(landscan) <- c("LandScan")
write_one(landscan, "landscan_on_landscan")
write_one(bimep_landscan_raster, "bimep_on_landscan")
```

## BIMEP Data on WorldPop Raster

The WorldPop pixel size is 100m, so it's (0.0008333,-0.0008333)
in lat-long.

```{r worlpop_raster}
worldpop <- raster::raster(fs::path(data_dir, "Equatorial_Guinea_100m_Population", "gnq_ppp_2018_UNadj.tif"))
worldpop <- raster::crop(worldpop, bimep_feature, snap = "out")
plot(worldpop)
```
Let's turn the worldpop into a 1km grid using aggregation.
```{r aggregate_worldpop}
worldpop_1km <- raster::aggregate(worldpop, fact = 10, fun = sum, expand = TRUE)
plot(worldpop_1km)
```
Now put the BIMEP on this raster.
```{r bimep_on_worldpop}
bimep_worldpop_raster <- raster::rasterize(
  bimep_feature,
  worldpop_1km,
  field = "pop",
  fun = sum
  )
names(bimep_worldpop_raster)
```

```{r}
plot(bimep_worldpop_raster)
```

```{r}
names(worldpop_1km) <- "WorldPop"
names(bimep_worldpop_raster) <- "BIMEP"
wp_layers <- raster::addLayer(worldpop_1km, bimep_worldpop_raster)
write_one(worldpop_1km, "worldpop_on_1kworldpop")
write_one(bimep_worldpop_raster, "bimep_on_1kworldpop")
write_one(worldpop, "worldpop_on_worldpop")
```
