---
title: "blurring"
output: html_notebook
---
# Overview

I want to find the average density of people within a one square kilometer
disc. I used a for-loop to sum over neighboring grid cells, but this is
both slow and imprecise, because I'd like to account for grid cells at
the edge of the circle. The speed matters a bit because I'd like to do this
10,000 times across a set of grids, in order to determine uncertainty from
the gridding process.

The `spatstat` package seems to have a really great function, but it
hasn't been working well. This notebook is me squirrelling around with
both the `spatstat::blur` command and modifications of it.

```{r}
library(raster)
library(spatstat)
```


```{r}
dims <- c(100, 100)  # row, column
range <- c(2000, 2000)  # width, height

dimension_coordinates <- function(range, dim) {
  nn <- 0:dim * range / dim
  0.5 * (nn[1:(length(nn) - 1)] + nn[2:length(nn)])
}

# xvals need columns, yvals need rows, so they are switched.
xvals <- dimension_coordinates(range[1], dims[2])
yvals <- dimension_coordinates(range[2], dims[1])
pixel_area <- (xvals[2] - xvals[1]) * (yvals[2] - yvals[1])
mat <- matrix(42.0, nrow = dims[1], ncol = dims[2])
mid <- dims %/% 2
mat[(mid[1] + 1):dims[1], 1:dims[2]] <- NA

fake_im <- spatstat::im(mat, xcol = xvals, yrow = yvals)
projected_raster <- raster::raster(maptools::as.SpatialGridDataFrame.im(fake_im))
```

```{r}
plot(projected_raster)
```


```{r}
Window(fake_im)
```
```{r}
plot(Window(fake_im))
```


```{r}
  sigma <- 0.5 * sqrt(10^6 / pi)
  raster_im <- maptools::as.im.RasterLayer(projected_raster)
  smoothed <- spatstat::blur(
    raster_im,
    sigma = sigma,
    kernel = "disc",
    normalise = FALSE,
    bleed = FALSE
    )
  smoothed_raster <- raster::raster(maptools::as.SpatialGridDataFrame.im(smoothed))
```

```{r}
vals <- as.array(smoothed)
c(min(vals, na.rm = TRUE), max(vals, na.rm = TRUE))
```
```{r}
plot(smoothed)
```


```{r}
plot(smoothed_raster)
```
```{r}
c(cellStats(smoothed_raster, stat = "min"), cellStats(smoothed_raster, stat = "max"))
```
```{r}
dims <- c(100, 100)  # row, column
range <- c(2000, 2000)  # width, height

dimension_coordinates <- function(range, dim) {
  nn <- 0:dim * range / dim
  0.5 * (nn[1:(length(nn) - 1)] + nn[2:length(nn)])
}

# xvals need columns, yvals need rows, so they are switched.
xvals <- dimension_coordinates(range[1], dims[2])
yvals <- dimension_coordinates(range[2], dims[1])
pixel_area <- (xvals[2] - xvals[1]) * (yvals[2] - yvals[1])
mat <- matrix(0.0, nrow = dims[1], ncol = dims[2])
mid <- dims %/% 2
mat[mid[1], mid[2]] <- 50
mat[(mid[1] + 1):dims[1], 1:dims[2]] <- NA

fake_im <- spatstat::im(mat, xcol = xvals, yrow = yvals)
projected_raster <- raster::raster(maptools::as.SpatialGridDataFrame.im(fake_im))
```


```{r}
  sigma <- 0.5 * sqrt(10^6 / pi)
  raster_im <- maptools::as.im.RasterLayer(projected_raster)
  smoothed <- spatstat::blur(
    raster_im,
    sigma = sigma,
    kernel = "disc",
    normalise = TRUE,
    bleed = FALSE
    )
  smoothed_raster <- raster::raster(maptools::as.SpatialGridDataFrame.im(smoothed))
```

```{r}
vals <- as.array(smoothed)
c(min(vals, na.rm = TRUE), max(vals, na.rm = TRUE))
```

```{r}
plot(smoothed)
```
```{r}
dims <- c(100, 100)  # row, column
range <- c(2000, 2000)  # width, height

dimension_coordinates <- function(range, dim) {
  nn <- 0:dim * range / dim
  0.5 * (nn[1:(length(nn) - 1)] + nn[2:length(nn)])
}

# xvals need columns, yvals need rows, so they are switched.
xvals <- dimension_coordinates(range[1], dims[2])
yvals <- dimension_coordinates(range[2], dims[1])
pixel_area <- (xvals[2] - xvals[1]) * (yvals[2] - yvals[1])
mat <- matrix(0.0, nrow = dims[1], ncol = dims[2])
mid <- dims %/% 2
peak <- 50
mat[mid[1], mid[2]] <- peak
mat[(mid[1] + 1):dims[1], 1:dims[2]] <- NA

# The factor of 2 is because that density is in a half circle.
max_density <- 2 * peak / 10^6
cat(paste("max density =", max_density, "\n"))

fake_im <- spatstat::im(mat, xcol = xvals, yrow = yvals)
projected_raster <- raster::raster(maptools::as.SpatialGridDataFrame.im(fake_im))
my_smoothed <- density_from_disc(projected_raster)
plot(my_smoothed)
```

