---
title: "Get Data"
output:
  pdf_document:
    number_sections: true
    df_print: kable
    toc: true
header-includes:
  - \usepackage{palatino}
# The params block means you can put the data anywhere you like and
# then render this document by passing a parameter to the rmarkdown render command.
params:
  data_directory: !r fs::path(rprojroot::find_root(rprojroot::is_rstudio_project), "inst", "extdata")
---

```{r knitr_options, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  autodep = TRUE,
  cache = TRUE
)
```


# Overview
This notebook retrieves input data and converts it into
gridded data about population density. This notebook should retrieve
and clean data so that we can pass that to another notebook to do statistics.

The two sections here regrid in two different ways. The first section
uses BIMEP GPS data, so we know where each house is. It re-grids
the GPS data to each gridded dataset with which we will compare.
You'll see that this matters for the 100m datasets but not for the 1km datasets.
The second section uses the BIMEP area and section grids. It realigns
HRSL, LandScan, and WorldPop to those grids.

## Concerns
I have an initial concern about population counts near
ocean borders. The population in the pixel will be entirely within the
land border, and this is handled differently by different datasets.
This could lead to different border population densities, and most
of the population is near the water.
We will see how much we can address this as we go.

We could also rescale populations according to the growth rate
for Bioko. A commonly-used adjustment is to
multiply values by $\exp(rt)$, where $r$ is the yearly growth rate.
This wouldn't affect the structure of zeroes, though.

## Conventions
Pixels in the water should be marked as NA and excluded from
calculations later.

All work will use pixels in latitude and longitude (lat-long). This
is also known as an unprojected space. We'll need to
transform the BIMEP projection to lat-long, but the rest
are already in lat-long, and the BIMEP coarse and fine grids
are rectilinear in lat-long.

We will ensure each dataset has NA values where it's ocean
and 0 values inland of the ocean.



```{r load_libraries, include = FALSE}
library(popbioko)
library(rprojroot)
library(sp)
library(sf)
library(tmap)
library(GISTools)
```
# Download Datasets
Specify a directory into which to put the data.
The `inst/extdata` subdirectory is a popular place.
```{r data_dir}
data_dir <- params$data_directory
if (!dir.exists(data_dir)) {
  dir.create(data_dir, recursive = TRUE)
}
```
The data will be stored in `r data_dir`.

The Bioko shapefile has one geometry, a polygon outline
of Bioko. We will use it to project population from rasters.
```{r predefinitions}
latitude_longitude_projection <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
bioko_sf <- sf::st_read(fs::path(data_dir, "source", "bioko.shp"))
```


If you have an account at IHME, the following will copy files to
your local drive, under the data directory. Look at the
help for `data_configuration` to see about creating a config file
for the download.
```{r download, results = "hide"}
download_worldpop(local_directory = data_dir)
download_bioko_grids(local_directory = data_dir)
download_hrsl_points(local_directory = data_dir)
```
# Align to Grids from BIMEP GPS Data
This section uses the original BIMEP GPS data,
so it's a longitude-latitude for each house an an integer
count of the number of people.
That means there is no section or area grid to consider.
We align BIMEP to any location.

The output of this section is a set of GeoTIFFS in the `aligned`
subdirectory of your data directory (likely `inst/extdata/aligned`).
The names are of the form `<grid><resolution>_<source>.tif`.
So the HRSL is on a 100m grid. It is saved as `HRSL100_HRSL.tif`.
Bioko data on that grid is saved as `HRSL100_Bioko.tif`.

For each grid, the code to project the GPS data is the same.
```{r project_bimep, code = capture.output(dump("bimep_on_grid", ""))}
```


## HRSL
The HRSL marks where houses are with values and everywhere else with NA.
The incoming data is for all of Equatorial Guinea, so we crop it.
We need to have zeroes where there is land, so add the zeroes
to the HRSL. Pixels with no land are marked NA.
Otherwise, we don't modify values.
```{r hrsl_original}
hrsl_raster <- read_hrsl(local_directory = data_dir)
hrsl_raster_crop <- raster::crop(hrsl_raster, bioko_sf, snap = "out")
hrsl_zero_mask <- raster::rasterize(bioko_sf, hrsl_raster_crop, field = 0)
hrsl_raster_zero <- raster::cover(hrsl_raster_crop, hrsl_zero_mask)
plot(hrsl_raster_zero)
```

Now put Bioko data on that grid:
```{r bioko_on_hrsl}
bimep_on_hrsl <- bimep_on_grid(hrsl_raster_zero, bioko_sf, local_directory = data_dir)
plot(bimep_on_hrsl)
```
```{r write_100m_hrsl}
popbioko::write_aligned_raster(
  hrsl_raster_zero,
  list(source = "HRSL", grid = "HRSL", resolution = 100),
  local_directory = data_dir
)
popbioko::write_aligned_raster(
  bimep_on_hrsl,
  list(source = "BIMEP", grid = "HRSL", resolution = 100),
  local_directory = data_dir
)
```


```{r write_1000m_hrsl}
aggregate_and_write <- function(raster, name) {
  aggregated <- raster::aggregate(raster, fact = 10, fun = sum, expand = TRUE)
  popbioko::write_aligned_raster(
    aggregated,
    list(source = name$source, grid = name$grid, resolution = 10 * name$resolution),
    local_directory = data_dir)
}
aggregate_and_write(
  hrsl_raster_zero,
  list(source = "HRSL", grid = "HRSL", resolution = 100)
)
aggregate_and_write(
  bimep_on_hrsl,
  list(source = "BIMEP", grid = "HRSL", resolution = 100)
)
```



## LandScan
LandScan is 1 km data, so we don't aggregate it.
```{r landscan_to_grid}
landscan <- read_landscan(local_directory = data_dir)
landscan <- raster::crop(landscan, bioko_sf, snap = "out")
bimep_on_landscan <- bimep_on_grid(
  landscan, bioko_sf, local_directory = data_dir)
popbioko::write_aligned_raster(
  landscan,
  list(source = "LandScan", grid = "LandScan", resolution = 1000),
  local_directory = data_dir
)
popbioko::write_aligned_raster(
  bimep_on_landscan,
  list(source = "BIMEP", grid = "LandScan", resolution = 1000),
  local_directory = data_dir
)
```
Actually, let's disaggregate it, to see what happens if we use it at 100m.
```{r disaggregate_landscan}
factor <- 10
landscan100 <- raster::disaggregate(landscan, fact = factor) / factor^2
bimep_on_landscan100 <- bimep_on_grid(
  landscan100, bioko_sf, local_directory = data_dir)
popbioko::write_aligned_raster(
  landscan100,
  list(source = "LandScan", grid = "LandScan", resolution = 100),
  local_directory = data_dir
)
popbioko::write_aligned_raster(
  bimep_on_landscan100,
  list(source = "BIMEP", grid = "LandScan", resolution = 100),
  local_directory = data_dir
)
```


## WorldPop

WorldPop is 100m, like the HRSL data, so we aggregate it.

```{r worldpop_to_grid}
worldpop <- raster::raster(fs::path(
  data_dir, "Equatorial_Guinea_100m_Population", "GNQ15v2.tif"))
worldpop <- raster::crop(worldpop, bioko_sf, snap = "out")
worldpop_id <- list(source = "WorldPop", grid = "WorldPop", resolution = 100)
bimep_id <- worldpop_id
bimep_id$source <- "BIMEP"
popbioko::write_aligned_raster(
  worldpop,
  worldpop_id,
  local_directory = data_dir
)
bimep_on_worldpop <- bimep_on_grid(
  worldpop, bioko_sf, local_directory = data_dir)
popbioko::write_aligned_raster(
  bimep_on_worldpop,
  bimep_id,
  local_directory = data_dir
)
aggregate_and_write(
  worldpop,
  worldpop_id
)
aggregate_and_write(
  bimep_on_worldpop,
  bimep_id
)
```

```{r plot_worldpop}
tm_shape(worldpop) + tm_raster()
```
```{r bimep_on_worldpop_plot}
tm_shape(bimep_on_worldpop) + tm_raster()
```

# Make other datasets align to BIMEP study grids
The BIMEP study uses two grids, at 100m and 1km resolution.
This section reads population datasets and regrids them
onto the BIMEP study datasets.


The output is all GeoTIFFs with names of the form data-on-grid.
The data are BIMEP, WorldPop, HRSL, and LandScan. The grids
are

- area, a 1km grid used by BIMEP.
- worldpop, a 100m grid used by WorldPop
- 1kworldpop, a 1km grid that aggregates and aligns with the WorldPop grid.
- landscan, the 1km LandScan grid
- hrsl, the 100m HRSL grid.

## Load BIMEP Early

The BIMEP data will have a "pop" feature for the population in the pixel.
Load it now so that we have its bounding box.

```{r bimep_raster}
bimep_feature <- bimep_population_as_points(local_directory = data_dir)
names(bimep_feature)
```
We see the population as `pop` and the other column, which is the geometry
for each point.

## Align HRSL and BIMEP to BIMEP 1k Grid
The Bioko grids define cells where BIMEP assigns residents of
the island. One is a 1 km grid, the other a 100 m grid.
If we read those grids, we can see that the input files
are projected into UTM 32N, which makes all distances in meters.
That's very convenient, but the original grids were defined as
a regular grid in latitude and longitude.
```{r grids, dependson = c("data_dir", "download")}
grids <- read_bioko_grids(local_directory = data_dir)
grid_info <- parameters_of_km_grid(grids$coarse, unproject = TRUE)
```
You can see from the description above that the coarse
and fine grids are shapefiles. Are these grids complete, or are they partial?
```{r complete_grids}
area_per_pixel <- function(grid) {
  bbox <- st_bbox(grid)
  (bbox["xmax"] - bbox["xmin"]) * (bbox["ymax"] - bbox["ymin"])
}
cat("area per pixel coarse", area_per_pixel(grids$coarse) / 3894, "\n")
cat("area per pixel fine", area_per_pixel(grids$fine) / 197086, "\n")
```
We see that the area per pixel for the coarse is about $10^6\:\mbox{m}^2$, which
means it's probably a complete grid. The fine grid looks far from complete
because we expect $10^4\:\mbox{m}^2$ per pixel.

If we want to make such a grid, we could use
```{r grid_copy, dpeendson = c("download")}
coarse_grid_copy <- make_fresh_grid(grid_info$bbox, grid_info$dimensions)
```
We can show the original and copy in unprojected space
in order to see that they overlap.
```{r compare_1km_grids, fig.cap = "The original and our copy overlap.", dependson = "grid_copy"}
grid_1km <- sf::st_transform(grids$coarse, crs = latitude_longitude_projection)
tm_shape(coarse_grid_copy) +
  tm_polygons(alpha = 0.1, border.col = "red") +
  tm_shape(grid_1km) +
  tm_polygons(alpha = 0.0, border.col = "blue")
```


What if we directly make a raster version of the 1km grid?
```{r make_raster}
km_raster <- raster::raster(
  ncol = grid_info$dimensions[1],
  nrow = grid_info$dimensions[2],
  xmn = grid_info$bbox["xmin"],
  xmx = grid_info$bbox["xmax"],
  ymn = grid_info$bbox["ymin"],
  ymx = grid_info$bbox["ymax"]
  )
raster::projection(km_raster) <- latitude_longitude_projection
```

Let's get some points to project to that raster.
The HRSL gives the location of a roof, or NA values. It doesn't say where
there could be population but isn't. Other population maps assign
a population of 0 to all non-ocean areas, including lakes. LandScan says
it extends the land area by a bit, too. We will use a shapefile for
Bioko in order to assign zero values to the HRSL.
```{r hrsl_plot}
hrsl_raster <- read_hrsl(local_directory = data_dir)
hrsl_raster_crop <- raster::crop(hrsl_raster, bimep_feature, snap = "out")
hrsl_zero_mask <- raster::rasterize(bioko_sf, hrsl_raster_crop, field = 0)
hrsl_raster_zero <- raster::cover(hrsl_raster_crop, hrsl_zero_mask)
hrsl_points_sf <- hrsl_points(hrsl_raster_crop)
# hrsl_points_sf[is.na(hrsl_points_sf[[1]]), 1] <- 0
plot(hrsl_raster_zero)
```


And do the projection using a builtin function of the raster package.
```{r project_to_raster}
km_hrsl_raster <- raster::rasterize(
  hrsl_points_sf,
  km_raster,
  field = "population_gnq_2018.10.01",
  fun = sum
  )
area_zero_raster <- raster::rasterize(bioko_sf, km_hrsl_raster, field = 0)
km_hrsl_raster <- raster::cover(km_hrsl_raster, area_zero_raster)
```

```{r}
plot(km_hrsl_raster)
```

Rasterize that pop feature, summing it per pixel.
```{r}
bimep_raster <- raster::rasterize(
  bimep_feature,
  km_raster,
  field = "pop",
  fun = sum
  )
names(bimep_raster)
```
```{r}
plot(bimep_raster)
```

We will write all raster layers the same way.
```{r}
write_one <- function(layer, basename) {
  raster::writeRaster(
    layer,
    filename = fs::path(data_dir, basename),
    format = "GTiff",
    overwrite = TRUE
    )
}
```

Write these as layers to a GeoTIFF.
```{r write_hrsl_bimep}
names(km_hrsl_raster) <- c("HRSL")
names(bimep_raster) <- c("BIMEP")
write_one(km_hrsl_raster, "hrsl_on_area")
write_one(bimep_raster, "bimep_on_area")
write_one(hrsl_raster_zero, "hrsl_on_hrsl")
```


```{r, eval = FALSE}
grids$coarse["hrsl"] <- hrsl_points_sf[[1]]
tm_shape(grids$coarse[grids$coarse$hrsl > 0,]) +
  tm_polygons("hrsl", title = "HRSL", palette = "Greens", style = "log10") +
  tm_layout(legend.title.size = 1, legend.text.size = 1)
```

## Align BIMEP to LandScan Grid

The LandScan grid is an Arc/Info Binary Grid, so it's the adf type.
It uses pixel sizes of 0.008333 in both dimensions.
It's the whole world, so we will subset it.

```{r landscan}
landscan <- read_landscan(local_directory = data_dir)
landscan <- raster::crop(landscan, bimep_feature, snap = "out")
plot(landscan)
```
Now put the bimep 100m data on the LandScan raster.
```{r}
bimep_landscan_raster <- raster::rasterize(
  bimep_feature,
  landscan,
  field = "pop",
  fun = sum
  )
names(bimep_landscan_raster)
```

```{r}
plot(bimep_landscan_raster)
```

```{r write_landscan_bimep}
names(bimep_landscan_raster) <- c("HRSL")
names(landscan) <- c("LandScan")
write_one(landscan, "landscan_on_landscan")
write_one(bimep_landscan_raster, "bimep_on_landscan")
```

## BIMEP Data on WorldPop Raster

The WorldPop pixel size is 100m, so it's (0.0008333,-0.0008333)
in lat-long.

```{r worlpop_raster}
worldpop <- raster::raster(fs::path(data_dir, "Equatorial_Guinea_100m_Population", "GNQ15v2.tif"))
worldpop <- raster::crop(worldpop, bimep_feature, snap = "out")
plot(worldpop)
```
Let's turn the worldpop into a 1km grid using aggregation.
```{r aggregate_worldpop}
worldpop_1km <- raster::aggregate(worldpop, fact = 10, fun = sum, expand = TRUE)
plot(worldpop_1km)
```
Now put the BIMEP on this raster.
```{r bimep_on_worldpop}
bimep_worldpop_raster <- raster::rasterize(
  bimep_feature,
  worldpop_1km,
  field = "pop",
  fun = sum
  )
names(bimep_worldpop_raster)
```

```{r}
plot(bimep_worldpop_raster)
```

```{r}
names(worldpop_1km) <- "WorldPop"
names(bimep_worldpop_raster) <- "BIMEP"
wp_layers <- raster::addLayer(worldpop_1km, bimep_worldpop_raster)
write_one(worldpop_1km, "worldpop_on_1kworldpop")
write_one(bimep_worldpop_raster, "bimep_on_1kworldpop")
write_one(worldpop, "worldpop_on_worldpop")
```

